\singlespacing \vspace{-1.6cm} \LARGE Trade-offs among plant reproductive traits determine interactions with floral visitors

### 2022-01-24 Manuscript/Lanuza_et_al_2021_Nat_Eco_Evo.Rmd
Editing for Ecol Letters

bibliography: 
  - references.bib
  - knitcitations.bib
# zotero: "My shared Zotero group"
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology
notes-after-punctuation: false
link-citations: yes
linkcolor: RoyalBlue
urlcolor: RoyalBlue
links-as-notes: false
subparagraph: yes

header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage[skip=3pt]{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{setspace}
- \usepackage{titlesec}
- \titlespacing{\title}{0pt}{\parskip}{-\parskip}
- \usepackage[style=nature]{biblatex}
- \usepackage[labelformat = empty]{caption}
- \usepackage{pdflscape}

@@

\vspace{-1.7cm}
\singlespacing
\large 
\textbf{Jose B. Lanuza$^{1,2}*$, Romina Rader$^{1}$, Jamie Stavert$^{3}$, Liam K. Kendall$^{4}$, Manu E. Saunders$^{1}$ and Ignasi Bartomeus$^{2}$}

\vspace{0.3cm}

\normalsize

__Plant life-history strategies are constrained by cost-benefit trade-offs that determine plant form and function. However, despite recent advances in the understanding of trade-offs for vegetative and physiological traits, little is known about plant reproductive economics and how they constrain plant life-history strategies and shape interactions with floral visitors. Here, we investigate plant reproductive trade-offs and how these drive interactions with floral visitors using a dataset of 17 reproductive traits for 1,506 plant species from 28 plant-pollinator studies across 18 countries. We tested whether a plant’s reproductive strategy predicts its interactions with floral visitors and if the different reproductive traits predict the plant’s role within the pollination network. We found that over half of all plant reproductive trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size, while the second axis indicated a pollinator dependency trade-off. Plant reproductive trade-offs helped explain partly the presence or absence of interactions with floral visitors, but not differences in visitation rate. However, we did find important differences in the interaction level among floral visitor guilds on the different axes of trait variation. Finally, we found that plant size and floral rewards were the most important traits in the understanding of the plant species network role. Our results highlight the importance of plant reproductive trade-offs in determining plant life-history strategies and plant-pollinator interactions in a global context.__

\small
\vspace{-0.5cm}

\noindent\rule{\textwidth}{1pt}

^1^ School of Environmental and Rural Science, University of New England, Armidale, New South Wales 2350, Australia. ^2^ Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain. ^3^ Department of Conservation | Te Papa Atawhai, Auckland, New Zealand. ^4^ Centre for Environmental and Climate Science, Lund University, Sölvegatan 37, S-223 62 Lund, Sweden. `*` e-mail: barragansljose@gmail.com

@@

Despite the astonishing diversity of floral structures among flowering plants [@barrett2002; @schiestl2013] and their importance in shaping plant-pollinator interactions [@fenster2004; @dellinger2020], a unified framework that explores plant reproductive trade-offs is currently lacking [@roddy2021]. In addition, macroecological studies that investigate plant reproductive traits are scarce [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017] and consequently, there is poor understanding of how reproductive traits drive interactions with floral visitors at large scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. Linking the plant's position in trait-space with the different pollinator groups could help to improve our understanding of plant-pollinator associations [@dehling2016]. Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016]. However, despite the generalist nature of most plant-pollinator interactions [@waser1996; @olesen2002], reproductive traits have been overlooked beyond highly specialised pollination systems [@dellinger2020]. Overall, it is unclear how specific plant reproductive biology traits shape plant-pollinator interactions [@tur2013; @devaux2014].

@@

Several studies have identified links between plant traits and plant-pollinator network properties [@bartomeus2013; @olito2015; @rowe2020]. Moreover, plant traits can define species’ network roles (e.g., specialists vs generalists) [@lazaro2013; @tur2013]. For example, plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators [@junker2013; @coux2016]. Morphological matching between plant and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure [@stang2009; @ibanez2012]. Remarkably, the combination of traits have shown to increase the predictive power of the network interactions [@eklof2013]. Therefore, considering the different plant reproductive trade-offs which represent the species reproductive strategy within the network [@dehling2016] could progress our understanding of plant-pollinator interactions. Further, we know little if those patterns generally studied at the community level are representative of wider macroecological scales.

>>>>

bibliography: 
  - references.bib
  - knitcitations.bib
# zotero: "My shared Zotero group"
csl: ecology.csl #ecology letters
notes-after-punctuation: false
link-citations: yes
linkcolor: RoyalBlue
urlcolor: RoyalBlue
links-as-notes: false
subparagraph: yes

header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage[skip=3pt]{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{setspace}
- \usepackage{titlesec}
- \titlespacing{\title}{0pt}{\parskip}{-\parskip}
- \usepackage[style=nature]{biblatex}
- \usepackage[labelformat = empty]{caption}
- \usepackage{pdflscape}
#- \DeclareUnicodeCharacter{0301}{*************************************}

@@

\vspace{-1.4cm}

\textbf{Jose B. Lanuza$^{1,2}$} <barragansljose@gmail.com>, \textbf{Romina Rader$^{1}$} <rrader@une.edu.au>, \textbf{Jamie Stavert$^{3}$} <jamie.stavert@gmail.com>, \textbf{Liam K. Kendall$^{4}$} <liam.k.kendall@gmail.com>, \textbf{Manu E. Saunders$^{1}$} <Manu.Saunders@une.edu.au> \textbf{and Ignasi Bartomeus$^{2}$} <nacho.bartomeus@gmail.com>

\small

^1^ School of Environmental and Rural Science, University of New England, Armidale, New South Wales 2350, Australia. ^2^ Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain. ^3^ Department of Conservation | Te Papa Atawhai, Auckland, New Zealand. ^4^ Centre for Environmental and Climate Science, Lund University, Sölvegatan 37, S-223 62 Lund, Sweden.

\doublespacing
\normalsize

\textbf{Statement of authorship:} JBL, RR and IB designed the study. JBL collated the data and conducted analysis with guidance of JS, LKK and IB. JBL wrote the manuscript with contributions of all authors.

\textbf{Data accessibility statement:} All data and code used to conduct this study will be deposited in the public repository Dryad upon acceptance.

\textbf{Short title:} Plant reproductive trade-offs.

\textbf{Keywords:} life-history strategies | plant reproductive trade-offs | plant-pollinator interactions.

\textbf{Type of article:} Letter.

\textbf{Number of words:} Abstract  (146 words), main text (XXXX).

\textbf{Number of references:} 77.

\textbf{Number of figures and tables:} 4 figures and 1 table.

\textbf{Corresponding author:} Jose B. Lanuza | +34 616375981| <barragansljose@gmail.com>

\newpage

# ABSTRACT

Plant life-history strategies are constrained by cost-benefit trade-offs that determine plant form and function. However, despite recent advances in the understanding of plant trait variation, little is known about plant reproductive trade-offs and how these constrain life-history strategies and shape interactions with floral visitors. Here, we investigate plant reproductive trade-offs and how these drive interactions with floral visitors using a dataset of 17 reproductive traits for 1,506 plant species. We found that over half of all plant reproductive trait variation was explained by two independent axes. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size, while the second axis indicated a pollinator dependency trade-off. In addition, plant reproductive trade-offs determined important differences in the interaction level among floral visitor guilds. Our study shows the main plant reproductive trade-offs and their relevance to understand plant-pollinator interactions in a global context.

\newpage

@@

**In bold PNAS editor's comments.**

- 1st explores is an ambiguous term. Maybe changing it for describe/determine? 2nd. Last phrase of the 1st paragraph lacks clarity. 3rd. Conceptual focus is unclear (3rd paragraph). General comment: What do we need to understand, and how will this information relate to broader problems in ecology/evolution? It is not sufficient to suggest that we need to improve our understanding without giving specifics regarding the impact of these studies.

- There is insufficient justification/discussion of the traits used (Table 1). For example, what is the expected relevance of plant height? In addition, some of the traits are classified as both quantitative and qualitative. How can this work? For example, the selfing level was given as four categories of quantitative estimates as well as four categories of qualitative estimates. Aren't these the same thing? Finally, it is not clear why some traits were included, such as ovule number, whole others were not, such as flower color.

- I am not convinced that the imputation process for missing trait values is accurate. There are many examples of very closely related plant species that differ dramatically in traits such as nectar volume and other traits relevant to pollinator interactions, so using a phylogenetic signal to estimate missing values may often give the wrong answer. In the Supplement you state: "In general terms, nectar data was rarely present and we tried to minimise the loss of information by filling some species at family level." Seems preferable to simply exclude nectar as one of your traits.

- A major issue with network data from studies of plant-visitor interactions is that there is rarely enough information to distinguish pollinators from pollen predators or nectar thieves.

- It is not clear that the broad phylogenetic scale investigated in the paper is appropriate for investigating the importance of trade-offs that evolve at the level of populations and species. I realize that the data at a finer micro-evolutionary scale are generally not available, but this is the level where constraints emerge.

# INTRODUCTION

Despite the astonishing diversity of floral structures among flowering plants [@barrett2002; @schiestl2013] and their importance in shaping plant-pollinator interactions [@fenster2004; @dellinger2020], a unified framework that **explores** plant reproductive trade-offs is currently lacking [@roddy2021]. In addition, macroecological studies that investigate plant reproductive traits are scarce [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017] and consequently, there is poor understanding of how reproductive traits drive interactions with floral visitors at large scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. Linking the plant's position in trait-space with the different pollinator groups could help to improve our understanding of plant-pollinator associations [@dehling2016]. Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016]. However, despite the generalist nature of most plant-pollinator interactions [@waser1996; @olesen2002], reproductive traits have been overlooked beyond highly specialised pollination systems [@dellinger2020]. **Overall, it is unclear how specific plant reproductive biology traits shape plant-pollinator interactions [@tur2013; @devaux2014].**

@@

Several studies have identified links between plant traits and plant-pollinator network properties [@bartomeus2013; @olito2015; @rowe2020]. Moreover, plant traits can define species’ network roles (e.g., specialists vs generalists) [@lazaro2013; @tur2013]. For example, plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators [@junker2013; @coux2016]. Morphological matching between plant and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure [@stang2009; @ibanez2012]. Remarkably, the combination of traits have shown to increase the predictive power of the network interactions [@eklof2013]. **Therefore, considering the different plant reproductive trade-offs which represent the species reproductive strategy within the network [@dehling2016] could progress our understanding of plant-pollinator interactions**. Further, we know little if those patterns generally studied at the community level are representative of wider macroecological scales.
